---
title: "<span style = 'font-size:14pt;'>**FWCE 457: ANOVA Class Notes </span>**"
author: "Emily Embury"
date: "`r Sys.Date()`"
output: pdf_document
---
<style type="text/css">
/* Whole document: */
body{
  font-size: 12pt;
}
/* Headers */
h1{
  font-size: 12pt;
}
</style>

---

```{r include = FALSE}
knitr::opts_chunk$set(
  echo = TRUE,              
  warning = FALSE,       
  message = FALSE,  
  cache = FALSE,        
  fig.align = "center", 
  fig.height = 4,
  fig.width = 5,
  class.source="bg-warning",
  class.output="bg-success"
)

```

# Analysis of variance (ANOVA)

 - Allows us to compare more than two independent groups (a paired t-test can only be used for two groups)
 - Essentially an extension of a two sample t-test
 - Types of ANOVA:
    - single-factor (one-way) ANOVA
    - multi-factor ANOVA
 - ANOVA design
    - predictor (or independent) variable: categorical
    - response (or dependent) variable: continuous
    

# One-way ANOVA
  - one categorical variable with more than 2 levels and a continuous response variable
  - each observation must be independent
  - terminology:
      - treatments: the different levels of the categorical variable that are being compared
      - replicates: observations made within each treatment
  - key assumptions:
      - observations are independent
      - normal distribution
      - variance is homogeneous (if variance is unequal = hetroscedasticity)
      - the residuals are normally distributed
      
      
```{r}
limp <- read.csv("limpet.csv", header = TRUE)
head(limp, 5) #density was converted to categorical variable for ANOVA
boxplot(limp$eggs ~ limp$density, xlab = "Density", ylab = "Number of egg masses") #numeric variable first then categorical
means <- tapply(limp$eggs, limp$density, mean)
points(means, col = "red", pch = 18)

#anova
mod <- aov(limp$eggs ~ limp$density)
summary(mod)

#calculate p-value
1 - pf(5.566, 3, 20)

```

# Post-hoc Analysis

  - conduct a pairwise analysis of the means
  - allows us to tell which means are different after rejecting null ANOVA
  - increases type 1 error
    - you can inflate your alpha by doing multiple comparisons with two-sample t-tests
    - using special post-hoc tests can help avoid alpha errors
  - tests
    - Tukey's HSD
    - Bonferroni
    - Scheffe
    
```{r}

#tukey HSD
TukeyHSD(mod)

```


# Check ANOVA assumptions
  - normal probability plot -- test normality
  - plot residuals -- test homogeneity of variance
  - plot residuals in time sequence -- test assumption of independence 
  
```{r}
par(mfrow = c(1,2)) #puts the plots next to each other
qqnorm(mod$residuals) #normal q-q plot, tests normality
qqline(mod$residuals, col = "red") #line of fit on q-q plot
plot(mod$fitted.values, mod$residuals, main = "Residuals vs Fitted", xlab = "Fitted Values", 
     ylab = "Residuals") #residuals vs. fitted, test variance
abline(h=0, col = "red") 
```
