---
title: "<span style = 'font-size:14pt;'>**FWCE 457: Exploratory Data Analysis Class Notes </span>**"
author: "Emily Embury"
date: "`r Sys.Date()`"
output: pdf_document
---
<style type="text/css">
/* Whole document: */
body{
  font-size: 12pt;
}
/* Headers */
h1{
  font-size: 12pt;
}
</style>

---

```{r include = FALSE}
knitr::opts_chunk$set(
  echo = TRUE,              
  warning = FALSE,       
  message = FALSE,  
  cache = FALSE,        
  fig.align = "center", 
  fig.height = 4,
  fig.width = 5,
  class.source="bg-warning",
  class.output="bg-success"
)

```

# **Data Types**

- Data can be classified based on the scale of measurement
  - Numerical data
      - Continuous (height, length, weight)
      - Discrete (number of offspring produced, number of animals captured)
  - Categorical data
     - Nominal (sex, habitat type, Occurrence (yes/no))
  - Ordinal (size (small, medium, large), age class (juvenile, adult))


- Graphs for numerical data
  - Histogram
  - Box plot
  - Scatter plot (for bivariate data)
- Graphs for categorical data
  - Bar plot
  - Pie chart


# **Graphical Representation**

#Example 1: Data on spine length (in mm) of 50 linyphiid spiders (source:Gotelli & Ellison (2013))

Histogram and Boxplot

```{r}
spider <- read.table("spider.txt", header = TRUE)

head(spider, n = 5) #view the first 5 observations

#histogram of spine length
hist(spider$length, xlab = "Spine length (mm)", main = "Histogram of spine length")

str(spider)
str(spider$length)


#boxplot of spine length
boxplot(spider$length, main = "Boxplot of spine length")
points(mean(spider$length), col = "red", pch = 18) #add mean (average) point
```


#Example 2: Data on abundance of species in different habitat types

- barplot and pie chart
- add two visualizations to the same figure

```{r}
habitat <- c("forest", "wetland")
abundance <- c(80, 60)
par(mfrow = c(1,2)) #allows you to put two plots in one
barplot(abundance, names.arg = habitat, ylab = "Abundance")
pie(abundance, labels = habitat)
```

#Example 3: Data on weight (g) and length (mm) of fish

- scatter plot

```{r}
weight <- c(300, 270, 1790, 360, 1150, 210, 660, 880, 550, 90)
length <- c(298, 280, 465, 315, 427, 254, 354, 394, 345, 162)

plot(length, weight, xlab= "Length (mm)", ylab = "Weight (g)")

length(weight) #check the number of counts, needs to be the same for scatter plots
length(length)
```


# **Descriptive Statistics**

# Measures of centeral tendency
- measures of location

- **Arithmatic Mean**
    - summation: add all possible values and divide by X
    - extreme values in dataset can skew mean

```{r}
spider <- read.table("spider.txt", header = TRUE)

mean(spider$length)
```

- **Median**
    - It divides the data into two halves with equal number of observations in
each half
    - is better when there are extreme values
    - to calculate:
        - arrange the data in increasing order
        - select the middle number
        - For an even number of observations, the median is simply the average of
the two middle numbers

```{r}
median(spider$length)
```
- **Mode**
    - Mode frequent value in the dataset
    - some data may not have a mode or may have more than one mode
    - helpful to describe the location of data especially with asymmetrical distribution

```{r}

```

- **Geometric mean**
    - can be helpful for measureing population growth
    - nth root of the product of observations
    - good for percentages
    
    - below calculation:
      - growth rate of owl pop. 10% and 20% over two years
      - average growth rate = below code = 14.89%

```{r}
growthr <- c(1.1, 1.2)
library(psych)
geometric.mean(growthr)
mean(growthr)
```

# **Measures of dispersion (variation)**

- measures of location not sufficient to describe data
- it is important to quantify the variability of observations


- **Variance**
  - measures variability of actual observed data
- Standard deviation
  - measures the square root of variance

```{r}
var(spider$length) #calculate variance
sd(spider$length) #calculate standard deviation
```

- **Quantiles**
  - Percentile is a kind of quantile
      - ex. if your test score is in 90th percentile, that means 90% of scores are below your score
      
      
  - median: 50th percentile
  - lower (1st) quartile is 25th percentile
  - upper (3rd) quartile is 75th percentile
```{r}
boxplot(spider$length, main = "Boxplot of spine length")
points(mean(spider$length), col = "red", pch = 18)

quantile(spider$length)
median(spider$length) #value is the same as 50th percentile
summary(spider$length)
```

- **Skewness**
  - Skewness describes how the sample shape differs from a symetrical distribution
  - it is a relationship between mean, median, and mode
  
  - **Symmetric** mean = median = mode
  - **Right-skewed** mean > median > mode
  - **Left-skewed** mean < median < mode