---
title: "<span style = 'font-size:14pt;'>**FWCE 457: Hypothesis Testing Class Notes </span>**"
author: "Emily Embury"
date: "`r Sys.Date()`"
output: pdf_document
---
<style type="text/css">
/* Whole document: */
body{
  font-size: 12pt;
}
/* Headers */
h1{
  font-size: 12pt;
}
</style>

---

```{r include = FALSE}
knitr::opts_chunk$set(
  echo = TRUE,              
  warning = FALSE,       
  message = FALSE,  
  cache = FALSE,        
  fig.align = "center", 
  fig.height = 4,
  fig.width = 5,
  class.source="bg-warning",
  class.output="bg-success"
)

```

# **Steps in hypothesis testing**
  1. state the null and alternate hypothesis
  2. specify level of significance
  3. collect data under one or more random sampling techniques
  4. choose an appropriate test statistic and its probability distribution
  5. determin the p-value
  6. make a decision to reject the null or fail to reject the null
  
  
# **Errors in hypothesis testing**

  - **type 1 error** -> if we reject the true null hypothesis
    - related to $\alpha$
  - **type 2 error** -> we fail to reject the null when it is false
    - related to $\beta$ 
    
    
# **One-sample t-test**
  - useful to mak inference about a single population mean
  - **Assumptions**
    - the population must be normally distributed
    - observations are independent
    
  - **Null hypothesis**
    - the population mean = a hypothesized value
    - always use = sign
    
  - **Alternative hypotheses**
    - the population mean $\ne$ a hypothesized value (two sided/two-tailed)
    - the population mean > a hypothesized value (one-sided/right-tailed)
    - the population mean < a hypothesized value (one-sided/left-tailed)
  
  - **Calculate p-value**
    - assume normal distribution, therefore p follows t-distribution table
    
  - **Example**
    - null: population mean of spine length = 0.27
    - alternative: population mean of spine length $\ne$ 0.27

```{r}

spider <- read.table("spider.txt", head = TRUE)

#summary statistics
summary(spider$length)
#right skewed, mean is greater than the median

#standard deviation
sd(spider$length)

#one-sample t-test
t.test(x = spider$length, alternative = "two.sided", mu = 0.27, conf.level = 0.95)

```
  - based on the **p-value**: since it is less than $\alpha$ = 0.05, we reject the null and conclude that the true mean spine length is different from 0.27mm
  - based on the **CI**: since the 95% CI does not include the true mean value (0.27mm), we reject the null
  
  
# **Paired t-test**
  - useful for when a measurement in one sample is matched to a sample in another sample (ex. before and after samples, recapturing birds)
  - **Assumptions**
    - the distribution of differences in paired data is normally distributed
    - the pairs are independent
  
  - **Null Hypothesis**
    - the population mean of the paired differences = 0

  - **Alternative hypotheses**
    - the population mean of the paired differences $\ne$ a hypothesized value (two sided/two-tailed)
    - the population mean of the paired differences > a hypothesized value (one-sided/right-tailed)
    - the population mean of the paired differences < a hypothesized value (one-sided/left-tailed)
  
  - **Example**
    - null: population mean of the paired body weight differences = 0
    - alternative: population mean of the paired body weight differences $\ne$ 0 
    - weight of the same individuals was measured in 2015 and 2016

```{r}
btit <- read.csv("bluetit.csv", header = TRUE)
head(btit, 3)
diff <- btit$weight2016 - btit$weight2015
summary(diff)
sd(diff)
meand <- mean(diff)
boxplot(diff, main = "Boxplot of body weight differences")
points(meand, col = "red", pch = 18)

t.test(x = btit$weight2016, y = btit$weight2015, paired = TRUE, alternative = "two.sided", 
       mu = 0, conf.level = 0.95, var.equal = TRUE)
```
  - based on the **p-value**: since it is more than $\alpha$ = 0.05, we fail to reject the null and conclude that there is no significant change in the mean weight of the blue tits
  - based on the **CI**: since the 95% CI does include 0, we fail to reject the null