---
title: "<span style = 'font-size:14pt;'>**FWCE 457: Multiple Linear Regression Class Notes </span>**"
author: "Emily Embury"
date: "`r Sys.Date()`"
output: pdf_document
---
<style type="text/css">
/* Whole document: */
body{
  font-size: 12pt;
}
/* Headers */
h1{
  font-size: 12pt;
}
</style>

---

```{r include = FALSE}
knitr::opts_chunk$set(
  echo = TRUE,              
  warning = FALSE,       
  message = FALSE,  
  cache = FALSE,        
  fig.align = "center", 
  fig.height = 4,
  fig.width = 5,
  class.source="bg-warning",
  class.output="bg-success"
)

```

# **Steps in multiple linear regression**

  - produce scatter plots to check whether there is a linear relationship between response and predictor variables
  - assess whether there is a multicollinearity problem
  - list and fit candidate models that relate the response variables to the predictor variables
  - conduct model selection
  - conduct residual diagnostics to assess the assumptions of the model and to identify outliers
  
  
# **Example**

```{r}
mayfly <- read.csv("mayfly.csv", header = TRUE)

plot(mayfly)
mcor <- cor(mayfly)
round(mcor, 2)
```

Multicollinearity
  - algae and BOD seem like they may be highly correlated