---
title: "<span style = 'font-size:14pt;'>**FWCE 457: Multiple Linear Regression Class Notes </span>**"
author: "Emily Embury"
date: "`r Sys.Date()`"
output: pdf_document
---
<style type="text/css">
/* Whole document: */
body{
  font-size: 12pt;
}
/* Headers */
h1{
  font-size: 12pt;
}
</style>

---

```{r include = FALSE}
knitr::opts_chunk$set(
  echo = TRUE,              
  warning = FALSE,       
  message = FALSE,  
  cache = FALSE,        
  fig.align = "center", 
  fig.height = 4,
  fig.width = 5,
  class.source="bg-warning",
  class.output="bg-success"
)

```

# **Steps in multiple linear regression**

  - produce scatter plots to check whether there is a linear relationship between response and predictor variables
  - assess whether there is a multicollinearity problem
  - list and fit candidate models that relate the response variables to the predictor variables
  - conduct model selection
  - conduct residual diagnostics to assess the assumptions of the model and to identify outliers
  
  
# **Example**

```{r}
mayfly <- read.csv("mayfly.csv", header = TRUE)

plot(mayfly)
mcor <- cor(mayfly)
round(mcor, 2)

```

Multicollinearity
  - algae and BOD seem like they may be highly correlated
  
  
#**Selecting the best model**
  - as # of parameters increase, the amount of variance increases
```{r}

M1 <- lm(Length ~ Speed, data = mayfly)
AIC(M1)
summary(M1)

M2 <- lm(Length ~ Algae, data = mayfly)
AIC(M2)
summary(M2)

M3 <- lm(Length ~ NO3, data = mayfly)
AIC(M3)
summary(M3)

M4 <- lm(Length ~ Speed + Algae, data = mayfly)
AIC(M4)
summary(M4)


M5 <- lm(Length ~ Speed + NO3, data = mayfly)
AIC(M5)
summary(M5)


M6 <- lm(Length ~ Algae + NO3, data = mayfly)
AIC(M6)
summary(M6)

#this model has the best R^2 and AIC value
M7 <- lm(Length ~ Speed + Algae + NO3, data = mayfly)
AIC(M7)
summary(M7)


#null control model
#an intercept on the model
M8 <- lm(Length ~ 1, data = mayfly)
AIC(M8)
summary(M8)

mean(mayfly$Length)



#plot residuals of best model to check assumptions
par(mfrow = c(1,2))
plot(M7, which = c(1, 2))
```