---
title: "<span style = 'font-size:14pt;'>**FWCE 457: Probability Distributions Class Notes </span>**"
author: "Emily Embury"
date: "`r Sys.Date()`"
output: pdf_document
---
<style type="text/css">
/* Whole document: */
body{
  font-size: 12pt;
}
/* Headers */
h1{
  font-size: 12pt;
}
</style>

---

```{r include = FALSE}
knitr::opts_chunk$set(
  echo = TRUE,              
  warning = FALSE,       
  message = FALSE,  
  cache = FALSE,        
  fig.align = "center", 
  fig.height = 4,
  fig.width = 5,
  class.source="bg-warning",
  class.output="bg-success"
)

```

# **Random Variables**

- a random variable is a variable that takes on different values, each with an associated probability
    - **Discrete Random Variables**
      - finite or countable values
        - presence or absence of a species
        - number of offspring produced
        
    - **Continuous Random Variables**
      - any value with a defined interval
        - biomass
        - wing length
        - pesticide concentration in fish tissues
        
# **Probability Distribution**

- obtained by assigning a probability to every possible outcome in an experiment
- **probabilities must be between 0 and 1**
  - the sum of a probability table must add up to 1
- two forms of probability distributions
  - discrete
    - Bernoulli, Binomial, Poisson, etc.
  - continuous
    - Uniform, Normal, Exponential
    
# **Bernoulli Distribution** (Discrete, two possible outcomes)
- *used for random variable with two possible outcomes*
- good for looking at **one** study site (n = 1)
- *Ex.* occurrence of a rare plant species at a site (Present or Not Present)
- can calculate mean and variance of the probability distribution

```{r}
rbinom(5, 1, 0.5) #like of flipping a coin 5 times, probability of observing a head (1) is 0.5
#rbinom calculates random values

a <- rbinom(10000, 1, 0.5)
table(a)

# detect a rare plant, the literature suggests that the probability of finding the plant is 2%
# find plant = 1
# do not find plant = 0

dbinom(x = 0, size = 1, prob = 0.02) # for Bernoulli, size = 1
#dbinom calculates distribution

# x = 0 because we want to calculate the probability of  not finding the plant or  X=0
# size = # of sites visited to find the plant

```
# **Binomial Distribution** (Discrete)
- good for looking across **multiple** study sites for species detection
- if you visit a site 5 times or visit 5 different sites, X can equal 0, 1, 2, 3, 4, or 5
- get a given probability for a single given site and apply to multiple sites
- like completing the Bernoulli distribution process multiple times

```{r}
# example: 
# number of detection of a rare plant in 25 site visits
# n = 25
# suppose the probability of detecting plant in one site is 0.02

dbinom(x = 0, size = 25, prob = 0.02) # probability of not seeing the species at all P(X = 0) in 25 sites
```
