---
title: "<span style = 'font-size:14pt;'>**FWCE 457: Probability Distributions Class Notes </span>**"
author: "Emily Embury"
date: "`r Sys.Date()`"
output: pdf_document
---
<style type="text/css">
/* Whole document: */
body{
  font-size: 12pt;
}
/* Headers */
h1{
  font-size: 12pt;
}
</style>

---

```{r include = FALSE}
knitr::opts_chunk$set(
  echo = TRUE,              
  warning = FALSE,       
  message = FALSE,  
  cache = FALSE,        
  fig.align = "center", 
  fig.height = 4,
  fig.width = 5,
  class.source="bg-warning",
  class.output="bg-success"
)

```

# **Random Variables**

- a random variable is a variable that takes on different values, each with an associated probability
    - **Discrete Random Variables**
      - finite or countable values
        - presence or absence of a species
        - number of offspring produced
        
    - **Continuous Random Variables**
      - any value with a defined interval
        - biomass
        - wing length
        - pesticide concentration in fish tissues
        
# **Probability Distribution**

- obtained by assigning a probability to every possible outcome in an experiment
- **probabilities must be between 0 and 1**
  - the sum of a probability table must add up to 1
- two forms of probability distributions
  - discrete
    - Bernoulli, Binomial, Poisson, etc.
  - continuous
    - Uniform, Normal, Exponential
    
# **Bernoulli Distribution** (Discrete, two possible outcomes)
- *used for random variable with two possible outcomes*
- good for looking at **one** study site or individual (n = 1)
- *Ex.* occurrence of a rare plant species at a site (Present or Not Present)
- can calculate mean and variance of the probability distribution

```{r}
rbinom(5, 1, 0.5) #like of flipping a coin 5 times, probability of observing a head (1) is 0.5
#rbinom calculates random values

a <- rbinom(10000, 1, 0.5)
table(a)

# detect a rare plant, the literature suggests that the probability of finding the plant is 2%
# find plant = 1
# do not find plant = 0

dbinom(x = 0, size = 1, prob = 0.02) # for Bernoulli, size = 1
#dbinom calculates distribution

# x = 0 because we want to calculate the probability of  not finding the plant or  X=0
# size = # of sites visited to find the plant

```
# **Binomial Distribution** (Discrete)
- good for looking across **multiple** study sites for species detection
- parameter of interest is probability
- if you visit a site 5 times or visit 5 different sites, X can equal 0, 1, 2, 3, 4, or 5
- get a given probability for a single given site and apply to multiple sites
- like completing the Bernoulli distribution process multiple times
- x >= 1 is "at least" one
- x <= 1 is "at most" one

```{r}
# example: 
# number of detection of a rare plant in 25 site visits
# n = 25
# suppose the probability of detecting plant in one site is 0.02

dbinom(x = 0, size = 25, prob = 0.02) #probability of not seeing the species at all P(X = 0) in 25 sites

dbinom(x= 1, size = 25, prob = 0.02) #probability of seeing the species at 1 out of 25 sites

### calculating with >= or =<

#probability of x =< 1
(dbinom(x = 1, size = 25, prob = 0.02) + dbinom(x = 0, size = 25, prob = 0.02))

#probability of x >= 1
  #P(x >= 1)
  # all probability add up to 1
  # 1 - P(x = 0)
1 - dbinom(x = 0, size = 25, prob = 0.02)

#pbinom vs. dbinom
  #dbinom : calculate P(X=x)
  #pbinom : calculate P(X =< x)
pbinom(q = 1, size = 25, prob = 0.02) # x is less than or equal to 1
dbinom(x = 1, size = 25, prob = 0.02) + dbinom(x = 0, size = 25, prob = 0.02) # x is less than or equal to 1
```
# **Poisson Distribution** (Discrete)
- possible values of X is some value x, no upper boundary
- parameter of interest is $\lambda$ where $\lambda$ is a the average # of occurrences of an event within a given time
- mean and variance are the same with Poisson distribution

```{r}
rpois(n = 5, lambda = 2.5) #random generation for the Poisson distribution with parameter lambda

# P(X>2) is very difficult because there is no upper boundary, 
 # need to subtract X = 0,1,2 from 1

1- ppois(q = 2, lambda = 0.75)

dpois(x = 3, lambda = 0.75) # probability that X = 3

ppois(q = 3, lambda = 0.75) # probability of x less than 3
```
# **Normal Distribution** (Continuous Probability Distribution)
- standard normal distribution: mean = 0 and variance = 1
- if you want to find the probability of X < 15, need to find the area under the curve < 15
```{r}
# ex. suppose the beetle size X is normally distributed with a mean 20mm and variance 10mm (st dev = 3.1623)
# calculate P(X<15)
pnorm(q = 15, mean = 20, sd = 3.1623)

# calculate P(X > 23.5) 
  # area under curve = 1 
  # there is no way to calculate area "above" in r so you need to subtract from 1
  # 1 - P(X<23.5)

1 - pnorm(q = 23.5, mean = 20, sd = 3.1623)

```

# ** Central Limit Theorem **
- a single population can be skewed but if you take multiple sets of samples, the distribution of those means will be a normal distribution
- sufficiently large sample size, you get a normal distribution