---
title: "<span style = 'font-size:14pt;'>**FWCE 457: Simple Linear Regression Class Notes </span>**"
author: "Emily Embury"
date: "`r Sys.Date()`"
output: pdf_document
---
<style type="text/css">
/* Whole document: */
body{
  font-size: 12pt;
}
/* Headers */
h1{
  font-size: 12pt;
}
</style>

---

```{r include = FALSE}
knitr::opts_chunk$set(
  echo = TRUE,              
  warning = FALSE,       
  message = FALSE,  
  cache = FALSE,        
  fig.align = "center", 
  fig.height = 4,
  fig.width = 5,
  class.source="bg-warning",
  class.output="bg-success"
)

```


# **Simple Linear Regression**

  - different from ANOVA because ANOVA requires categorical variables
  - regression analysis compares two or more numeric variables
  - can be used to asses the strength of the relationship and make predictions
  - scatterplot is best to assess data
  - pearson correlation coefficient
    - indicates the **linear** relationship between groups
    - indicates strong and weak relationships
    - will not show anything about a quadratic relationship
    - correlation does not imply causation
    - value = 0.7 or -0.7 -- strong
    - value = 0.5-0.7 or -0.7-(-0.5) -- medium
    
# **Pearson Correlation Coefficient**
```{r}
mayfly <- read.csv("mayfly.csv", header = TRUE)
head(mayfly, 4)
plot(mayfly$Algae, mayfly$Length, xlab = "Algae cover (%)", ylab = "Length (mm)")

cor(mayfly$Algae, mayfly$Length) #correlation coefficient
  # 0.7650757 indicates a strong positive linear relationship (over .7 = strong)
```

# **Linear Regression Model **
  - first variable must be response variable
  - second variable is predictor variable
  
```{r}
m1 <- lm(mayfly$Length ~ mayfly$Algae)
summary(m1)

plot(mayfly$Algae, mayfly$Length, xlab = "Algae cover (%)", ylab = "Length (mm)")
lines(mayfly$Algae, m1$fitted.values, col = "red", lwd = 2)

```

  - to get the predicted values for algae cover %, plug the % into the fitted regression line equation as X
  - 12.566 + 0.121 * X
  - 40% algae cover
    - 12.566 + 0.121 * 40 = 17.406
    - if you use m1$fitted.values, this number matches the fitted values for 40% algae cover
    
    
# ** Measuring degree of fit**
  - how well does the regression line fit the data?
  - high $R^2$ indicates a better fit
  
```{r}
summary.aov(m1)
```
# **Check assumptions**    
  - assume normal distribution and constant variance
  
```{r}

par(mfrow = c(1,2))
plot(m1, which = c(1,2))

```